---
title: "Assignment 04: Tidy data and joins"
author: "Almas K."
date: '2019-10-07'
output:
  html_document:
    keep_md: TRUE
---

---
```{r load, warning=FALSE,echo=FALSE}
suppressPackageStartupMessages(library(tidyverse)) 
suppressPackageStartupMessages(library(gapminder))
```

# Exercise 1: Univariate Data Reshaping

## Univariate Option 1:Life Expectancy vs Year

### 1.1 Data in Wider Format
Let's choose 2 other countries to compare to Canada 
```{r}
gapwide <- gapminder %>% 
  filter((country=="Algeria")|(country=="Bahrain")|(country=="Canada")) %>%
  pivot_wider(id_cols= year, 
              names_from= country,
              values_from = lifeExp) 
  gapwide %>%
    DT::datatable()
```

### 1.2 Scatter Plot for 2 countries against each other:

Let's choose Algeria and Bahrain to plot against each other.
```{r}
gapwide %>%
  ggplot(aes(x=Algeria, y=Bahrain))+
  geom_point()+
  labs(title="Scatter Plot of Bahrain's Life Expectancy vs Algeria's Life Expectancy",x="Algeria(Years)",y="Bahrain(Years)")+
  theme_bw()
```

Over time, life Expectancy of Bahrain is higher compared to Algeria 

### 1.3 Relengthen Data:
```{r}
gapwide %>%
  pivot_longer(cols = c(-year),
               names_to = "Country",
               values_to="Life Expectancy") %>%
  DT::datatable()
               
```

# Excercise 2: Multivariate Data Reshaping:

## Multivariate Option 1: Life Expectancy-Gdp

### 2.1 Wide Tibble
```{r}
gap_mulit_wide <- gapminder %>%
filter(country== "Australia"| country == "Bahrain") %>%
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = c(lifeExp,gdpPercap))
gap_mulit_wide %>% knitr::kable()
```

### 2.2 Relengthen Data:
```{r}
gap_mulit_wide %>%
  pivot_longer(cols=c(-year),
               names_to = c(".value", "country"),
               names_sep = "_") %>%
  knitr::kable()

## .value separates values for the measures of lifeExp and gdpPerCap
```

# Excercise 3.0 Table Joins:

### Loading the guest and email data
```{r}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```
###  3.01 First Edit email csv:
Each guest should on one line for each email and guest field changed to name to make it easier to join

```{r}
Email <- email %>%
  separate_rows(email,guest, sep = ",") %>%
  select(name=guest,email) %>%
 mutate(name=trimws(name, "both"))

## There is a white space in Email dataset, removed this 
```

## 3.1 Add Emails for Each Guest on List:

Assumption: Only including guests from guest list who have a corresponding email from the email tibble: 
 (no additonal emails, no additional guests):
```{r}
guest_emails <- Email %>%
  inner_join(guest,Email, by="name")

guest_emails %>% DT::datatable()
```

## 3.2 Emails but not on Guest List

The names of peoples with emails but are not on the guest list:

```{r}
guest_emails %>%
  select(name,email) %>%
  setdiff(Email,.) %>%
  knitr::kable()
```

These 3 are not found in the guest dataset. 

## 3.3 Guest List for all those that we have emails for:

A new guest list with names of people on the original guest list with emails and those not on the original guest list who also have emails (found on email dataset but not guest dataset)

```{r}
Email %>%
  left_join(guest,Email, by="name") %>%
DT::datatable() 
```

 

### 3--Extra Analysis:
There are 5 guests who don't have emails in the email dataset.  These are the ones that don't have a corresponding email in the email tibble but are found on the guest list (guest dataset)

Let's find out who the five guests on the guest list that don't have emails are : 
```{r}
guest_emails%>%
  select(-email) %>%
  setdiff(guest,.) %>%
  knitr::kable()
```


