---
title: "Assignment 04: Tidy data and joins"
author: "Almas K."
date: '2019-10-07'
output:
  html_document:
    keep_md: TRUE
---

---
```{r load, warning=FALSE,echo=FALSE}
suppressPackageStartupMessages(library(tidyverse)) 
suppressPackageStartupMessages(library(gapminder))
```

# Exercise 1: Univariate Data Reshaping

## Univariate Option 1:Life Expectancy vs Year

### 1.1 Data in Wider Format
Let's choose 2 other countries to compare to Canada 
```{r}
gapwide <- gapminder %>% 
  filter((country=="Algeria")|(country=="Bahrain")|(country=="Canada")) %>%
  pivot_wider(id_cols= year, 
              names_from= country,
              values_from = lifeExp) %>%
print(gapwide) 
  
```
### 1.2 Scatter Plot for 2 countries against each other:

Let's choose Algeria and Bahrain to plot against each other.
```{r}
gapwide %>%
  ggplot(aes(x=Algeria, y=Bahrain))+
  geom_point()+
  theme_bw()
```

Over time, life Expectancy of Bahrain is higher compared to Algeria 

### 1.3 Relengthen Data:
```{r}
gapwide %>%
  pivot_longer(cols = c(-year),
               names_to = "Country",
               values_to="Life Expectancy")
               
```

# Excercise 2: Multivariate Data Reshaping:

## Multivariate Option 1: Life Expectancy-Gdp

### 2.1 Wide Tibble
```{r}
gap_mulit_wide <- gapminder %>%
filter(country== "Australia"| country == "Bahrain") %>%
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = c(lifeExp,gdpPercap))
print(gap_mulit_wide) 
```

### 2.2 Relengthen Data:
```{r}
gap_mulit_wide %>%
  pivot_longer(cols=c(-year),
               names_to = c(".value", "country"),
               names_sep = "_")

## .value separates values for the measures of lifeExp and gdpPerCap
```

# Excercise 3.0 Table Joins:

### Loading the guest and email data
```{r}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```
###  3.1.1 First Edit email csv:
Each guest should on one line for each email and guest field changed to name to make it easier to join

```{r}
Email <- email %>%
  separate_rows(email,guest, sep = ",") %>%
  select(name=guest,email) %>%
 mutate(name=trimws(name, "both"))

## There is a white space in Email dataset, removed this 
```

## 3.1 Add Emails for Each Guest on List:

Assumption: Only including guests from guest list who have a corresponding email from the email tibble: 
 (no additonal emails, no additional guests):
```{r}
guest_emails <- Email %>%
  inner_join(guest,Email, by="name") %>%
print(guest_emails)
```

## 3.2 Emails but not on Guest List

The names of peoples with emails but are not on the guest list:

```{r}
guest_emails %>%
  select(name,email) %>%
  setdiff(Email,.)
```

## 3.3 Guest List for all those that we have emails for:

A new guest list with names of people on the original guest list with emails and those not on the original guest list who also have emails 

```{r}
guest_all_emails <- Email %>%
  left_join(guest,Email, by="name") %>%
print(guest_emails)
```

There are 5 guests who don't have emails in the email tibble but are on the guest list. 

###Extra Analysis:

Let's find out who the five guests on the guest list that don't have emails are : 
```{r}
guest_emails%>%
  select(-email) %>%
  setdiff(guest,.)
```
These are the ones that don't have a corresponding email in the email tibble. 
