---
title: "Hw02-Explore Gapminder and use dplyr"
author: "Almas K."
date: '2019-09-20'
output: github_document
---

```{r load, warning=FALSE,echo=FALSE}
suppressPackageStartupMessages(library(tidyverse)) 
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(knitr))
```
# Excercise 1:

## 1.1 Filter for selected 3 countries in the 1970s 
```{r}
kable(
  gapminder %>% 
  filter(year >= 1970 & year <= 1979, country == "Canada"|country == "Sweden"|country == "Pakistan"))
```

## 1.2 Select country and gdpPercap from 1.1 dataset
```{r}
kable(
gapminder %>% 
filter(year >= 1970 & year <= 1979, country == "Canada"|country == "Sweden"|country == "Pakistan") %>%
select(country, gdpPercap))
```

## 1.3 Filter gapminder to all entries that have experienced a drop in life expectancy.
```{r}
gapminder %>%
  group_by(country) %>%
mutate(increaseLifeExp=lifeExp-lag(lifeExp)) %>%
  filter(increaseLifeExp<0) 
```

## 1.4 Filter gapminder so that it shows the max GDP for each country:
```{r}
gapminder %>% 
group_by(country) %>%
  summarize(max(gdpPercap))
```

## 1.5 Scatterplot of Canada's Life Expentancy vs GDP per Capita
```{r}
gapminder %>%
  filter(country == "Canada") %>%
  ggplot(aes(log(gdpPercap),lifeExp))+
  geom_point()
```

## Excercise 2: : Explore individual variables with dplyr
We will look at continent as our categorical and population as our quantitative in the gapminder dataset.

### 2.1 : Summary Table
```{r}
kable(
gapminder %>% 
  group_by(continent) %>% #This tells you the possible values for continent
 summarize(min(pop) # The minimum population by each continent
           ,max(pop) # The maximum (min and max make up range)
           ,mean(pop) #Th mean population
           ,sd(pop), #The standard deviation or the spread of the population
           IQR(pop))) #The Interquartile range (also another measure of spread)
```

From this we can see 5 possible values for continent, as well as the range and spread. Now onto the distribution for #2:

### 2.2 Density Plot 
```{r}
gapminder %>% 
  ggplot(aes(x=log(pop)))+ #Transform to log 
  geom_density()+ #Plot density
    facet_wrap(. ~continent) #Allows you to plot the density plots separately for each continent and side by side 

```

The population was log transformed. After log transformation, it is interesting to note the bimodal distribution for Oceania and Europe's population data

## Excercise 3 : Produce Various Plots
I will be using a the ____ dataset from the R datasets package, since we are allowed to use this package. This data deals with

## Plot 1: 
