---
title: "Hw02-Explore Gapminder and use dplyr"
author: "Almas K."
date: '2019-09-20'
output: github_document
---

```{r load, warning=FALSE,echo=FALSE}
library(tidyverse) 
library(gapminder)
```
# Excercise 1:

##1.1 Filter for selected 3 countries in the 1970s 
```{r}
gapminder %>% 
  filter(year >= 1970 & year <= 1979, country == "Canada"|country == "Sweden"|country == "Pakistan")
```
## 1.2 Select country and gdpPercap from 1.1 dataset
```{r}
gapminder %>% 
  filter(year >= 1970 & year <= 1979, country == "Canada"|country == "Sweden"|country == "Pakistan") %>%
select(country, gdpPercap)
```
## 1.3 Filter gapminder to all entries that have experienced a drop in life expectancy.
```{r}
gapminder %>%
  group_by(country) %>%
mutate(increaseLifeExp=lifeExp-lag(lifeExp)) %>%
  filter(increaseLifeExp<0) 
```
## 1.4 Filter gapminder so that it shows the max GDP for each country:
```{r}
gapminder %>% 
group_by(country) %>%
  summarize(max(gdpPercap))

gapminder %>%
 arrange(desc(gdpPercap)) %>%
 slice(1:3,1702:1704)

aggregate(gapminder$gdpPercap, by=list(gapminder$country), max)
```
## 1.5 Scatterplot of Canada's Life Expentancy vs GDP per Capita
```{r}
gapminder %>%
  filter(country == "Canada") %>%
  ggplot(aes(log(gdpPercap),lifeExp))+
  geom_point()
```
##Excercise 2: : Explore individual variables with dplyr
We will look at continent as our categorical and population as our quantitative in the gapminder dataset
```{r}
gapminder %>% 
  group_by(continent) %>% #This tells you the possible values for continent
 summarize(min(pop) # The minimum population by each continent
           ,max(pop) # The maximum (min and max make up range)
           ,mean(pop) #Th mean population
           ,sd(pop), #The sd or the spread of the population
           IQR(pop)) #The Interquartile range (also part of spread)
```
From this we can see 5 possible values for continent, as well as the range and spread. Now onto the density and 
## Excercise 3 : Produce Various Plots

