---
title: "Hw02-Explore Gapminder and use dplyr"
author: "Almas K."
date: '2019-09-20'
output: github_document
---

```{r load, warning=FALSE,echo=FALSE}
suppressPackageStartupMessages(library(tidyverse)) 
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(datasets))
```
# Excercise 1:

## 1.1 Filter for selected 3 countries in the 1970s 
```{r}
kable(
  gapminder %>% 
  filter(year >= 1970 & year <= 1979, country == "Canada"|country == "Sweden"|country == "Pakistan"))
```

## 1.2 Select country and gdpPercap from 1.1 dataset
```{r}
kable(
gapminder %>% 
filter(year >= 1970 & year <= 1979, country == "Canada"|country == "Sweden"|country == "Pakistan") %>%
select(country, gdpPercap))
```

## 1.3 Filter gapminder to all entries that have experienced a drop in life expectancy.
```{r}
  gapminder %>%
  group_by(country) %>%
mutate(increaseLifeExp=lifeExp-lag(lifeExp)) %>%
  filter(increaseLifeExp<0)
```

## 1.4 Filter gapminder so that it shows the max GDP for each country:
```{r}
gapminder %>% 
group_by(country) %>%
  summarize(max(gdpPercap))
```

## 1.5 Scatterplot of Canada's Life Expentancy vs GDP per Capita
```{r}
gapminder %>%
  filter(country == "Canada") %>%
  ggplot(aes(log(gdpPercap),lifeExp))+
  geom_point()
```

## Excercise 2: : Explore individual variables with dplyr
We will look at continent as our categorical and population as our quantitative in the gapminder dataset.

### 2.1 : Summary Table
```{r}
kable(
gapminder %>% 
  group_by(continent) %>% #This tells you the possible values for continent
 summarize(min(pop) # The minimum population by each continent
           ,max(pop) # The maximum (min and max make up range)
           ,mean(pop) #Th mean population
           ,sd(pop), #The standard deviation or the spread of the population
           IQR(pop))) #The Interquartile range (also another measure of spread)
```

From this we can see 5 possible values for continent, as well as the range and spread. Now onto the distribution for #2:

### 2.2 Density Plot 
```{r}
gapminder %>% 
  ggplot(aes(x=log(pop)))+ #Transform to log 
  geom_density()+ #Plot density
    facet_wrap(. ~continent) #Allows you to plot the density plots separately for each continent and side by side 

```

The population was log transformed. After log transformation, it is interesting to note the bimodal distribution for Oceania and Europe's population data

## Excercise 3 : Produce Various Plots
I will be using mtcars dataset from the R datasets package for Plot 1. This data deals with different aspects of vehicle design for 32 cars.

## Plot 1: Scatter Plot of Horsepower vs Miles per Gallon for Automatic Transmissions
```{r}
mtcars %>% 
  filter(am==0) %>% ##Filter for automatics transmissions
  ggplot(aes(x=mpg,y=hp)) + 
  geom_point()+
  labs(x="Miles per Gallon",y="Horsepower")


```
There is an overall negative association between miles per gallon and horsepower for automatic transmission cars here. 
##Plot 2: Bar Graph Showing for each cylinder and transmission type 
```{r}
mtcars %>%
  mutate(cyl=as.factor(cyl)) %>% #Change cylinder to Factor
  mutate(automatic_transmiss=if_else(am==0,"Yes","No")) %>% # created new factor for transmission
  ggplot(aes(x=cyl,group=automatic_transmiss))+ #Chose aesthetics
  geom_bar(aes(fill=automatic_transmiss))+ #Added colour fill for bar chart 
  labs(x="Number of cylinders") #changed x labels 

```
Most cars in the dataset were 8 cylinder , and most of the 8 cylinder were automatic transmission.
## Recycling:
```{r}
filter(gapminder, country == c("Rwanda", "Afghanistan")) 

gapminder %>% 
  filter(country =="Rwanda" | country =="Afghanistan")
```
No the analyst misssed about 12 datapoints. The reason is because they used == and a column bind, which relies on both sets of data to be the same size,thus it recycles the c(Afghanistan,Rwanda) until it is the length of the longest vector, which in this case is  
#
