---
title: "Hw03_dplyr_ggplot2_Part_II"
author: "Almas K."
date: '2019-09-28'
output: github_document
---
```{r load, warning=FALSE,echo=FALSE}
suppressPackageStartupMessages(library(tidyverse)) 
suppressPackageStartupMessages(library(gapminder))
```

## Task Option 1:Report the absolute and/or relative abundance of countries with low life expectancy over time by continent: Compute some measure of worldwide life expectancy – you decide – a mean or median or some other quantile or perhaps your current age. Then determine how many countries on each continent have a life expectancy less than this benchmark, for each year.
```{r}
gapminder %>%
  group_by(year) %>%
  mutate(median_lifeExp=median(lifeExp)) %>%
  ungroup(year) %>%
  mutate(less_than_median= if_else(lifeExp<median_lifeExp,TRUE,FALSE)) %>%
  filter (less_than_median==TRUE) %>%
  group_by(continent,year,less_than_median) %>%
  summarize(n_less_than_median=sum(less_than_median)) %>%
  print() %>%
  ggplot(aes(x=year, y=n_less_than_median,group=continent))+
  geom_bar(aes(fill=continent),position="dodge",stat="identity")

gapminder %>%
  group_by(year) %>%
  mutate(median_lifeExp=median(lifeExp)) %>%
  ungroup(year) %>%
  mutate(less_than_median= if_else(lifeExp<median_lifeExp,TRUE,FALSE)) %>%
  group_by(continent,year,less_than_median) %>%
  count()
  
```



## Task Option 3 : Look at the spread of GDP per capita within the continents.
```{r}
gp_spread <- gapminder %>%
   group_by(continent)  %>%
mutate(log_gdpPercap= log(gdpPercap)) %>%
  arrange(log_gdpPercap)
gp_spread %>%
  summarize(min(gdpPercap),
            max(gdpPercap),
            sd(gdpPercap),
            IQR(gdpPercap))
gp_spread %>%
  ggplot(aes(x=log_gdpPercap)) +
  geom_density() +
   facet_wrap(. ~continent) 


```

THe distribution is right skewed for Africa, left-skewed for Europe, mostly normal for the Americas and Oceania, and It is interesting to note that the Americas are the

## Task Option 4: Compute a trimmed mean of life expectancy for different years. Or a weighted mean, weighting by population. Just try something other than the plain vanilla mean.
```{r}
gapminder %>%
   group_by(continent,year)  %>%
  summarize(wt_mean=weighted.mean(lifeExp,pop,na.rm = TRUE)) %>%
  print() %>%
  ggplot(aes(x=year,y=wt_mean)) +
  geom_point()+
  facet_wrap(. ~continent) 
```



