---
title: "Homework 1_Gapminder Exploration Presentation"
output: ioslides_presentation
---

```{r load, warning=FALSE,echo=FALSE}
library(tibble) 
library(gapminder)
```
## A basic look at the gapminder rows and columns {.smaller}
```{r}
gpmd <- gapminder
 ncol(gpmd) # number of columns 
 nrow(gpmd) # number of rows 
colnames(gpmd) # column names 
```
## Let's learn more about the structure of the data {.smaller}
```{r}
str(gpmd,strict.width="wrap") #basic summary of data structure
```
## The range of years included in the data are:
```{r}
range(gpmd$year) 
```
## How many countries are there in the dataset:
```{r}
sum(nlevels(gpmd$country))
```
# Let's explore life-expectancies from the dataset a bit:

## What are the ranges of life expectancies?
```{r}
range(gpmd$lifeExp)
```
## Which countries have the highest and lowest life-expectancies?
```{r}
gpmd$country[max(gpmd$lifeExp)] #highest
gpmd$country[min(gpmd$lifeExp)] #lowest
```

## Average Life expectancy per continent:
```{r}
aggregate(lifeExp~continent,gpmd,mean) #includes all years
```
## Let's learn about the median GDP per capita for each continent
```{r}
aggregate(gdpPercap~continent,gpmd,median) #includes all years
```
